---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  word_document: default
  html_notebook: default
  pdf_document: default
---


(a) Use read.delim function to read Su_raw_matrix.txt into a variable called su (Notice that su has become a data frame now):

```{r}
# Su_raw_matrix.txt is in my working directory
su <- read.delim("Su_raw_matrix.txt")

```

(b) Use mean and sd functions to find mean and standard deviation of Liver_2.CEL column:

```{r}
# "Liver_2.CEL"
mean_value <- mean(su$Liver_2.CEL)
sd_value <- sd(su$Liver_2.CEL)

# Print the results
cat("Mean of Liver_2.CEL:", mean_value, "\n")
cat("Standard Deviation of Liver_2.CEL:", sd_value, "\n")

```
(c) Use colMeans and colSums functions to get the average and total values of each column:

```{r}
# su is the data frame
average_values <- colMeans(su)
total_values <- colSums(su)

# Print the results
cat("Average values of each column:\n")
print(average_values)

cat("\nTotal values of each column:\n")
print(total_values)

```
## QUESTION 2
Use rnorm(n, mean = 0, sd = 1) function in R to generate 10000 numbers for the following (mean, sigma) pairs
and plot histogram for each, meaning you need to change the function parameter accordingly. Then comment on how
these histograms are different from each other and state the reason. (20 points)
(a) mean=0, sigma=0.2
(b) mean=0, sigma=0.5

```{r}
# Set seed for reproducibility
set.seed(123)

# Generate 10000 random numbers for each case
data_a <- rnorm(10000, mean = 0, sd = 0.2)
data_b <- rnorm(10000, mean = 0, sd = 0.5)

# Plot histograms
par(mfrow = c(2, 1))  # Arrange plots in 2 rows, 1 column

hist(data_a, main = "Histogram for (a) mean=0, sigma=0.2", col = "lightblue", breaks = 30)
hist(data_b, main = "Histogram for (b) mean=0, sigma=0.5", col = "lightgreen", breaks = 30)

```

##Question 3
```{r}
# (a) Create the 'dat' dataframe
library(ggplot2)
dat <- data.frame(cond = factor(rep(c("A", "B"), each = 200)),
                  rating = c(rnorm(200), rnorm(200, mean = 0.8)))
```


```{r}
# (b) Overlaid histograms
ggplot(dat, aes(x = rating, fill = cond)) +
  geom_histogram(binwidth = 0.5, alpha = 0.5, position = "identity")
```

```{r}
# (c) Interleaved histograms
ggplot(dat, aes(x = rating, fill = cond)) +
  geom_histogram(binwidth = 0.5, position = "dodge")
```


```{r}
# (d) Density plots
ggplot(dat, aes(x = rating, colour = cond)) +
  geom_density()
```


```{r}
# (e) Density plots with semitransparent fill
ggplot(dat, aes(x = rating, fill = cond)) +
  geom_density(alpha = 0.3)
```


```{r}
# (f) Read "diabetes_train.csv" into a variable called diabetes
diabetes <- read.csv("diabetes_train.csv")

# Overlaid histograms for 'mass' attribute of diabetes
ggplot(diabetes, aes(x = mass, fill = class)) +
  geom_histogram(binwidth = 0.5, alpha = 0.5, position = "identity")
```


```{r}
# Interleaved histograms for 'mass' attribute of diabetes
ggplot(diabetes, aes(x = mass, fill = class)) +
  geom_histogram(binwidth = 0.5, position = "dodge")


```


```{r}
# Density plots for 'mass' attribute of diabetes
ggplot(diabetes, aes(x = mass, colour = class)) +
  geom_density()

```


```{r}

# Density plots with semitransparent fill for 'mass' attribute of diabetes
ggplot(diabetes, aes(x = mass, fill = class)) +
  geom_density(alpha = 0.3)
```
#A
Explanation: This operation is performed using the %>% (pipe) operator from the dplyr package.
drop_na(): Removes rows with missing values (NA) from the dataset.
summary(): Provides a summary of the dataset, including statistical measures for each column.
```{r}
# Load the necessary packages
library(tidyverse)

# Read the titanic.csv file from DATA folder to a variable named passengers
passengers <- read.csv("titanic.csv")

# (a) Remove rows with missing values and provide a summary of the dataset
summary_data <- passengers %>% drop_na() %>% summary()
print(summary_data)

```
##B
Explanation: This operation filters the dataset to include only rows where the 'Sex' column is equal to "male".
filter(Sex == "male"): Retains only those rows where the 'Sex' column is "male".

```{r}

# (b) Filter the data to include only male passengers
male_passengers <- passengers %>% filter(Sex == "male")
print(male_passengers)
```
#C
Explanation: This operation arranges the dataset in descending order based on the 'Fare' column.
arrange(desc(Fare)): Sorts the dataset in descending order of 'Fare'.

```{r}
# (c) Arrange the data in descending order based on Fare
sorted_passengers <- passengers %>% arrange(desc(Fare))
print(sorted_passengers)

```
#D
Explanation: This operation adds a new column 'FamSize' to the dataset, representing the sum of 'Parch' (number of parents/children aboard) and 'SibSp' (number of siblings/spouses aboard).
mutate(FamSize = Parch + SibSp): Creates a new column 'FamSize' with the sum of 'Parch' and 'SibSp'. This can help in analyzing the family size of each passenger.

```{r}
# (d) Create a new column 'FamSize' that is the sum of 'Parch' and 'SibSp'
passengers_with_famsize <- passengers %>% mutate(FamSize = Parch + SibSp)
print(passengers_with_famsize)
```

#E
Explanation: This operation groups the dataset by the 'Sex' column and calculates summary statistics within each group.
group_by(Sex): Groups the data by the 'Sex' column, allowing for separate calculations for male and female passengers.
summarise(meanFare = mean(Fare), numSurv = sum(Survived)): Calculates the mean 'Fare' and total number of 'Survived' passengers for each gender group. This can provide insights into average fare expenses and survival rates based on gender.
```{r}
# (e) 
grouped_passengers <- passengers %>% group_by(Sex) %>% summarise(meanFare = mean(Fare), numSurv = sum(Survived))
print(grouped_passengers)
```

#Question 6
```{r}
percentiles <- quantile(diabetes$skin, c(0.1, 0.3, 0.5, 0.6))

# Print the results
cat("10th percentile:", percentiles[1], "\n")
cat("30th percentile:", percentiles[2], "\n")
cat("50th percentile (median):", percentiles[3], "\n")
cat("60th percentile:", percentiles[4], "\n")
```


```{r}
```

